---
title: "Adding analyses to a plan"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Adding analyses to a plan}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  chunk_output_type: console
---

This vignette builds three plans on `nor_covid19_cases_by_time_location`, a dataset of Covid-19 cases in Norway. `vignette("plnr")` defines the terms: argset, action function, analysis and plan.

## Single-function plan

A single-function plan applies one action function to many argsets. Add the argsets first. Then give every argset the same action function with `apply_action_fn_to_all_argsets()`.

### Many strata

This plan draws one graph for each location.

```{r, collapse=FALSE}
library(ggplot2)
library(data.table)

p <- plnr::Plan$new()

data_fn <- function() {
  return(plnr::nor_covid19_cases_by_time_location)
}
p$add_data(name = "covid19_cases", fn_name = "data_fn")

location_codes <- unique(p$get_data()$covid19_cases$location_code)
location_codes

p$add_argset_from_list(
  plnr::expand_list(
    location_code = location_codes,
    granularity_time = "isoweek"
  )
)
p$get_argsets_as_dt()

action_fn <- function(data, argset) {
  if (plnr::is_run_directly()) {
    data <- p$get_data()
    argset <- p$get_argset(1)
  }
  pd <- data$covid19_cases[
    location_code == argset$location_code &
      granularity_time == argset$granularity_time
  ]

  q <- ggplot(pd, aes(x = date, y = covid19_cases_testdate_n))
  q <- q + geom_line()
  q <- q + labs(title = argset$location_code)
  q
}

p$apply_action_fn_to_all_argsets(fn_name = "action_fn")

q <- p$run_all()
q[[1]]
q[[2]]
```

### Many variables

This plan crosses two choices: cases or cases per 100 000 population, and weekly or daily data. It draws one graph for each of the four combinations.

```{r, collapse=FALSE}
p <- plnr::Plan$new()

data_fn <- function() {
  return(plnr::nor_covid19_cases_by_time_location[location_code == "nation_nor"])
}
p$add_data(name = "covid19_cases", fn_name = "data_fn")

p$add_argset_from_list(
  plnr::expand_list(
    variable = c("covid19_cases_testdate_n", "covid19_cases_testdate_pr100000"),
    granularity_time = c("isoweek", "day")
  )
)
p$get_argsets_as_dt()

action_fn <- function(data, argset) {
  if (plnr::is_run_directly()) {
    data <- p$get_data()
    argset <- p$get_argset(1)
  }
  pd <- data$covid19_cases[granularity_time == argset$granularity_time]

  q <- ggplot(pd, aes(x = date, y = .data[[argset$variable]]))
  q <- q + geom_line()
  q <- q + labs(title = paste(argset$variable, argset$granularity_time))
  q
}

p$apply_action_fn_to_all_argsets(fn_name = "action_fn")

p$run_one(1)
p$run_one(2)
p$run_one(3)
p$run_one(4)
```

## Multi-function plan

A multi-function plan gives each analysis its own action function. This plan makes the figures of a short report. Figure 1 has its own action function, and figures 2 and 3 share one.

```{r, collapse=FALSE}
p <- plnr::Plan$new()

data_fn <- function() {
  return(plnr::nor_covid19_cases_by_time_location)
}
p$add_data(name = "covid19_cases", fn_name = "data_fn")

figure_1 <- function(data, argset) {
  if (plnr::is_run_directly()) {
    data <- p$get_data()
    argset <- p$get_argset("figure_1")
  }
  pd <- data$covid19_cases[granularity_time == "isoweek"]

  q <- ggplot(pd, aes(x = date, y = covid19_cases_testdate_pr100000))
  q <- q + geom_line()
  q <- q + facet_wrap(~location_code)
  q <- q + labs(title = "Weekly Covid-19 cases per 100 000 population")
  q
}

plot_epicurve_by_location <- function(data, argset) {
  if (plnr::is_run_directly()) {
    data <- p$get_data()
    argset <- p$get_argset("figure_2")
  }
  pd <- data$covid19_cases[
    granularity_time == "isoweek" &
      location_code == argset$location_code
  ]

  q <- ggplot(pd, aes(x = date, y = covid19_cases_testdate_n))
  q <- q + geom_line()
  q <- q + labs(title = argset$location_code)
  q
}

p$add_analysis(name = "figure_1", fn_name = "figure_1")
p$add_analysis(
  name = "figure_2",
  fn_name = "plot_epicurve_by_location",
  location_code = "nation_nor"
)
p$add_analysis(
  name = "figure_3",
  fn_name = "plot_epicurve_by_location",
  location_code = "county_nor03"
)

p$run_one("figure_1")
p$run_one("figure_2")
p$run_one("figure_3")
```
