---
title: "Introduction to plnr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to plnr}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

# Introduction

`plnr` is a framework for planning and executing analyses in R. Use it to organize and run multiple analyses. It covers two cases: the same function applied with different arguments, and multiple different functions applied to your data.

## Core Concepts

### Broad technical terms

+----------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Object**                             | **Description**                                                                                                                                                     |
+----------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| argset                                 | A named list containing a set of arguments.                                                                                                                         |
+----------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| analysis                               | These are the fundamental units that are scheduled in `plnr`:                                                                                                       |
|                                        |                                                                                                                                                                     |
|                                        | -   1 argset                                                                                                                                                        |
|                                        | -   1 (action) function that takes two arguments                                                                                                                    |
|                                        |     1.  data (named list)                                                                                                                                           |
|                                        |     2.  argset (named list)                                                                                                                                         |
+----------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| plan                                   | This is the overarching "scheduler":                                                                                                                                |
|                                        |                                                                                                                                                                     |
|                                        | -   1 data pull                                                                                                                                                     |
|                                        | -   1 list of analyses                                                                                                                                              |
+----------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------+

### Different types of plans

+----------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Plan Type**                          | **Description**                                                                                                                                                     |
+----------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Single-function plan                   | Same action function applied multiple times with different argsets applied to the same datasets.                                                                    |
+----------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Multi-function plan                    | Different action functions applied to the same datasets.                                                                                                            |
+----------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------+

### Plan Examples

+----------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Plan Type**                          | **Example**                                                                                                                                                         |
+----------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Single-function plan                   | Multiple strata (e.g. locations, age groups) that you need to apply the same function to to (e.g. outbreak detection, trend detection, graphing).                   |
+----------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Single-function plan                   | Multiple variables (e.g. multiple outcomes, multiple exposures) that you need to apply the same statistical methods to (e.g. regression models, correlation plots). |
+----------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Multi-function plan                    | Creating the output for a report (e.g. multiple different tables and graphs).                                                                                       |
+----------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------+

## Basic Usage

This simple example shows the core concepts:

```{r, collapse=FALSE}
library(plnr)
library(ggplot2)
library(data.table)

# Create a new plan
p <- Plan$new()

# Add data
p$add_data(
  name = "deaths",
  direct = data.table(deaths=1:4, year=2001:2004)
)

# Add argsets for different years
p$add_argset(
  name = "fig_1_2002",
  year_max = 2002
)

p$add_argset(
  name = "fig_1_2003",
  year_max = 2003
)

# Define analysis function
fn_fig_1 <- function(data, argset) {
  plot_data <- data$deaths[year <= argset$year_max]
  
  ggplot(plot_data, aes(x=year, y=deaths)) +
    geom_line() +
    geom_point(size=3) +
    labs(title = glue::glue("Deaths from 2001 until {argset$year_max}"))
}

# Apply function to all argsets
p$apply_action_fn_to_all_argsets(fn_name = "fn_fig_1")

# Run analyses
p$run_one("fig_1_2002")
```

## Advanced Features

### Data Management

The framework makes data management efficient in three ways:

- It loads data once and reuses it across analyses.
- It keeps data cleaning separate from analysis.
- It tracks data changes with a hash.

### Debugging Tools

`plnr` includes several tools for development and debugging:

```{r, collapse=FALSE}
# Access data directly
p$get_data()

# Access specific argset
p$get_argset("fig_1_2002")

# Access analysis by name or index
p$get_analysis(1)

# Use is_run_directly() for development
fn_analysis <- function(data, argset) {
  if(plnr::is_run_directly()) {
    data <- p$get_data()
    argset <- p$get_argset("fig_1_2002")
  }
  
  # function continues here
}
```

### Function Naming

When you add an analysis, you can use either `fn_name` or `fn`:

```{r, collapse=FALSE}
# Using fn_name (recommended)
p$add_analysis(
  name = "fig_1_2002",
  fn_name = "fn_fig_1",
  year_max = 2002
)

# Using fn (for function factories)
p$add_analysis(
  name = "fig_1_2003",
  fn = fn_fig_1,
  year_max = 2003
)
```

### Hash-based Caching

The framework uses hashing to track data changes:

```{r, collapse=FALSE}
# Create two plans with same data
p1 <- Plan$new()
p1$add_data(direct = data.table(deaths=1:4, year=2001:2004), name = "deaths")
p1$add_data(direct = data.table(deaths=1:4, year=2001:2004), name = "deaths2")

p2 <- Plan$new()
p2$add_data(direct = data.table(deaths=1:4, year=2001:2004), name = "deaths")
p2$add_data(direct = data.table(deaths=1:4, year=2001:2004), name = "deaths2")

# Same data has same hash
identical(p1$get_data()$hash$current_elements, p2$get_data()$hash$current_elements)

# Different data has different hash
p1$add_data(direct = data.table(deaths=1:5, year=2001:2005), name = "deaths3")
p1$get_data()$hash$current_elements
```

## Best Practices

1.  **Data Organization**
    -   Keep data cleaning separate from analysis.
    -   Use meaningful names for datasets.
    -   Document the data structure and the assumptions.
2.  **Analysis Functions**
    -   An action function MUST accept `data` and `argset` parameters.
    -   Use `is_run_directly()` during development.
    -   Keep each function focused on one purpose.
3.  **Plan Structure**
    -   Use meaningful names for argsets and analyses.
    -   Group related analyses together.
    -   Document the plan structure and its dependencies.
4.  **Development Workflow**
    -   Start with small examples.
    -   Use the debugging tools during development.
    -   Test each analysis on its own, before you run the full plan.

## Next Steps

1.  Read the [Adding Analyses](https://www.rwhite.no/plnr/articles/adding_analyses.html) vignette for more detailed examples.
2.  Visit the [package website](https://www.rwhite.no/plnr/) for additional resources.
3.  Read the function documentation with `help(package="plnr")`.
